Imputation of incomplete multilevel data

### 2024-03-13 manuscript.qmd
analyze imputed data

All incomplete variable have a multilevel imputation method.

@@

@@

<!-- 1.	Define the most general analytic model to be applied to imputed data; -->
<!-- 2.	Select a 2l method that imputes close to the data; -->
<!-- 3.	Include all level-1 variables; -->
<!-- 4.	Include the disaggregated cluster means of all level-1 variables; -->
<!-- 5.	Include all level-1 interactions implied by the analytic model; -->
<!-- 6.	Include all level-2 predictors; -->
<!-- 7.	Include all level-2 interactions implied by the analytic model; -->
<!-- 8.	Include all cross-level interactions implied by the analytic model; -->
<!-- 9.	Include predictors related to the missingness and the target; -->
<!-- 10.	Exclude any terms involving the target -->

<!-- 1. Include all variables that appear in the analysis model, including transformation, interactions and outcomes -->
<!-- 2. Include all variables that are related to the nonresponse -->
<!-- 3. Include all variables that explain a considerable amount of variance -->
<!-- 4. Remove variables that have too many missing values within the subgroup of incomplete cases -->

---

## Sandbox

Fix cluster-level missingness determinately.
```{r}
# meth <- make.method(dat)
# meth["experience_j"] <- "2lonly.mean"
# meth[meth != "2lonly.mean"] <- ""
# pred <- make.predictorMatrix(dat)
# pred[] <- 0
# pred[, "cluster_id"] <- -2
# imp <- mice(
#   popular_MAR, 
#   m = 1, 
#   method = meth, 
#   predictorMatrix = pred,
#   maxit = 1)
# plot_pattern(complete(imp))
```

This procedure has fixed the missing experience value

Does not work if there are inconsistencies within clusters! But we can do this manually.

```{r}
# dat |>
#   group_by(cluster_id) |>
#   reframe(teacher_experience = unique(experience_j))
```

This shows us that in the first cluster teacher experience of 4 and 24 is recorded.

```{r}
# dat |>
#   filter(cluster_id == 1) |>
#   ggplot(aes(experience_j)) + geom_histogram()
```

For most pupils, teacher experience is 24. We can quite safely assume that the teacher experience of 4 is a data entry error. We can fix this by over-imputing the teacher experience of 4 to 24, and deductively imputing the missing teacher experience from the same cluster as 24. The following line of code sets the teacher experience of 4 manually as missing.

```{r}
# dat[dat$cluster_id == 1 & (is.na(dat$experience_j) | dat$experience_j == 4), "experience_j"] <- NA
```

```{r}
# imp <- mice(
#   popular_MAR, 
#   m = 1, 
#   method = meth, 
#   predictorMatrix = pred,
#   maxit = 1)
# plot_pattern(complete(imp))
```

>>>>

All incomplete variables have a multilevel imputation method.

@@

# Visualize

```{r}
ggmice(imp, aes(y = popularity_ij, x = extraversion_ij, group = cluster_id)) +
  geom_point(alpha = 0.1) +
  geom_smooth(color = "grey", method = "lm", se = FALSE, alpha = 0.1) +
  geom_smooth(aes(group = NULL), color = "black", method = "lm", se = FALSE, alpha = 0.1)
```

@@

#### Single-level model

```{r, echo=TRUE}
cca <- lm(
  popularity_ij ~ 1, 
  data = dat
  )
tidy(cca)
```

```{r, echo=TRUE}
fit <- with(
  imp,
  lm(popularity_ij ~ 1, data = dat)
  ) |> 
  pool()
tidy(fit)
```

#### Intercept-only model
```{r}
cca <- lmer(
  popularity_ij ~ (1 | cluster_id),
  data = dat,
  REML = FALSE
  )
tidy(cca)
```

```{r, echo=TRUE}
fit <- with(
  imp,
  lmer(
    popularity_ij ~ (1 | cluster_id), 
    data = dat, 
    REML = FALSE)
  ) 
pool(fit) |> tidy()
```

```{r, echo=TRUE}
mitml::testEstimates(as.mitml.result(fit), extra.pars = TRUE)
```

#### Model with explanatory variables

```{r}
cca <- lmer(
  popularity_ij ~ gender_ij + extraversion_ij + experience_j + (1 | cluster_id),
  data = dat,
  REML = FALSE
  )
tidy(cca)
```

```{r, echo=TRUE}
fit <- with(
  imp,
  lmer(
    popularity_ij ~ gender_ij + extraversion_ij + experience_j + (1 | cluster_id),
    data = dat, 
    REML = FALSE)
  ) 
pool(fit) |> tidy()
```

```{r, echo=TRUE}
mitml::testEstimates(as.mitml.result(fit), extra.pars = TRUE)
```

#### Model with explanatory variables, extraversion slope random

```{r}
cca <- lmer(
  popularity_ij ~ gender_ij + extraversion_ij + experience_j +
    (1  + extraversion_ij | cluster_id),
  data = dat,
  REML = FALSE
  )
tidy(cca)
```

```{r, echo=TRUE}
fit <- with(
  imp,
  lmer(
    popularity_ij ~ gender_ij + extraversion_ij + experience_j + 
      (1  + extraversion_ij | cluster_id), 
    data = dat, 
    REML = FALSE)
  ) 
pool(fit) |> tidy()
```

```{r, echo=TRUE}
mitml::testEstimates(as.mitml.result(fit), extra.pars = TRUE)
```

#### Model with with cross-level interaction

```{r}
cca <- lmer(
  popularity_ij ~ gender_ij + extraversion_ij + experience_j +
    extraversion_ij:experience_j + (1  + extraversion_ij | cluster_id),
  data = dat,
  REML = FALSE
  )
tidy(cca)
```

```{r, echo=TRUE}
fit <- with(
  imp,
  lmer(
    popularity_ij ~ gender_ij + extraversion_ij + experience_j + 
      extraversion_ij:experience_j + (1  + extraversion_ij | cluster_id),  
    data = dat, 
    REML = FALSE)
  ) 
pool(fit) |> tidy()
```

```{r, echo=TRUE}
mitml::testEstimates(as.mitml.result(fit), extra.pars = TRUE)
```
